---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Louise Nyholm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 4

In this assignment we do the following:
- we run a Bayesian meta-analysis of pitch variability in ASD, based on previously published literature
- we analyze pitch variability in ASD in two new studies using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Perform a meta-analysis of pitch variability from previous studies of voice in ASD
- the data is available as Ass4_MetaAnalysisData.tsv
- You should calculate Effect size (cohen's d) and Standard Error (uncertainty in the Cohen's d) per each study, using escalc() from the metafor package (also check the livecoding intro)
- N.B. we're only interested in getting a meta-analytic effect size for the meta-analytic prior (and not e.g. all the stuff on publication bias). See a brms tutorial here: https://vuorre.netlify.com/post/2016/09/29/meta-analysis-is-a-special-case-of-bayesian-multilevel-modeling/ The formula is EffectSize | se(StandardError) ~ 1 + (1 | Paper). Don't forget prior definition, model checking, etc.
- Write down the results of the meta-analysis in terms of a prior for step 2.

```{r}
# Loading packages
library(pacman)
pacman::p_load(
  tidyverse, 
  metafor,
  brms) 

# Loading data
MA_d <- read_delim("Ass4_MetaAnalysisData.tsv", delim = "\t")

# Tidying data (making sure the relevant variables are categorised correctly)
MA_d <- MA_d %>%
  mutate(
    PitchVariabilityASD_Mean = as.numeric(PitchVariabilityASD_Mean),
    PitchVariabilityTD_Mean = as.numeric(PitchVariabilityTD_Mean),
    PitchVariabilityASD_SD = as.numeric(PitchVariabilityASD_SD),
    PitchVariabilityTD_SD = as.numeric(PitchVariabilityTD_SD)
  )

# Only keeping the studies with a name (the NA rows have no data all)
MA_d <- MA_d %>%
  subset(!is.na(Paper))

# Using escalc() to calculate Effect size (cohen's d) and Standard Error (uncertainty in the Cohen's d) per each study
MA_d <- escalc(measure = "SMD", # Standardized mean difference
            n1i = TD_N, # Specifying group size of TD
            n2i = ASD_N, # Specifying group size of ASD
            m1i = PitchVariabilityTD_Mean, # Specifying mean of TD
            m2i = PitchVariabilityASD_Mean, # Specifying mean of ASD
            sd1i = PitchVariabilityTD_SD, # Specidying  SD of TD
            sd2i = PitchVariabilityASD_SD, # Specifying SD of ASD
            data = MA_d, # DATA
            slab = Paper) # (Optional) - labels for the studies

# Renaming yi (effect size) and calcultting SE from vi (variance)
MA_d <- MA_d %>% 
  mutate(
    StandardError = sqrt(vi) # Why is this not the SD (vs. SE) - SD = within sample "error", SE = between sample "error" (og det er meta analyse så automatisk between studies i guess?)
    ) %>%
  rename(
  EffectSize = yi
  )

# Looking at summary of the effect sizes and the standard errors
summary(MA_d$EffectSize)
summary(MA_d$StandardError)
# Should we take into consideration that different measures are used? Maybe not include the ones with SDs and semitones as measures of variability? And only include Hz (most)? Or does it not matter, because we use varying effects for the studies/populations? Then just remember that it has an effect when "simplifying" it like here in a summary with a mean and min + max.

# Specifying a formula
#we can use either Paper or Population, because their values correspond - when the paper name is the same (indicated by last names and years), the population value will also be the same (indicated by letters) - so they should be able to be used interchangably.
MA_f <- bf(EffectSize | se(StandardError) ~ 1 + (1 | Paper))

# Defining priors
#get prior for at finde ud af hvilke priors vi skal definere.
get_prior(MA_f, data = MA_d, family = gaussian())



MA_prior <- c(
  prior(normal(0, 1), class = Intercept), 
  prior(normal(0, .3), class = sd)
)

# Prior predictive check
MA_m0 <- brm(
  MA_f,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = "only",
  chains = 2,
  cores = 2,
  file = "MA_m0"
)

pp_check(MA_m0, nsamples = 100)
# Men dataen ser ud til at være binomial?

# Fitting the model
MA_m1 <- brm(
  MA_f,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = T,
  chains = 2,
  cores = 2,
  file = "MA_m1"
)

#hvis du opdaterer modellerne, så husk at slet filerne, ellers bliver de ikke opdateret

# Posterior predictive check
pp_check(MA_m1, nsamples = 100)

# Looking at the estimates
summary(MA_m1)
# RESULTS: MA effect mean = 0.43, sd = 0.1 #vs. mean = -0.43, SD = 0.09

# Saving the results in variables to use later
MA_mean <- fixef(MA_m1)[[1]] # Defining the effect size of intercept as the mean
MA_se <- fixef(MA_m1)[[2]] # Defining the SD as mean_se (WHY SE?)

MA_heterogeneity = 0.32 # Defining the sd(Intercept) (group-level effects) as heterogeneity

```



Step 2: Analyse pitch variability in ASD in two new studies for which you have access to all the trials (not just study level estimates)
- the data is available as Ass4_data.csv. Notice there are 2 studies (language us, and language dk), multiple trials per participant, and a few different ways to measure pitch variability (if in doubt, focus on pitch IQR, interquartile range of the log of fundamental frequency (hvilket vi tror er HZ?))
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).
- Is there any structure in the dataset that we should account for with random/varying effects? How would you implement that? Or, if you don't know how to do bayesian random/varying effects or don't want to bother, is there anything we would need to simplify in the dataset?

```{r}
d <- read_csv("Ass4_data.csv", col_types = cols(ID = col_character()))

d <- d %>% mutate(
  PitchVariability = scale(Pitch_IQR)
)

mean(d$PitchVariability) #=0 fordi scaling? også mean centered?

hist(d$Pitch_IQR)
hist(d$PitchVariability)

```



Step 3: Build a regression model predicting Pitch variability from Diagnosis.
- how is the outcome distributed? (likelihood function). NB. given we are standardizing, and the meta-analysis is on that scale, gaussian is not a bad assumption. Lognormal would require us to convert the prior to that scale.
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on? Start simple, with Diagnosis only. Add other predictors only if you have the time and energy!
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Evaluate model quality. Describe and plot the estimates. 

```{r}

NewStudies_f0 <- bf(PitchVariability ~ 1 + Diagnosis + (1|ID)) #skrives med 1 for at få ændringen fra TD til ASD? Nej omvendt, fordi ASD er vores "baseline" forid det starter med A.
NewStudies_f1 <- bf(PitchVariability ~ 0 + Language + Language:Diagnosis + (1|ID))

summary(d$PitchVariability)
get_prior(NewStudies_f0, d, family = gaussian())

NS_prior0 <- c(
  prior(normal(0, .3), class = Intercept), 
  prior(normal(0, .1), class = b), # because it is a 1+ model the beta is the difference between the diagnosis. we are scpetical, therefor we expect a mean beta coeficcient of 0. #fordi vi har scale det --> mean = 0. og SE = 0.1 fordi vi forventer små effekter i psykologiske "eksperimenter"
  prior(normal(0, .1), class = sd), # we expect different people to be just as different from each other as the average person with autism is different from the average TD person. We dont expect the variability between people to be different from the variability between 
  prior(normal(.5, .3), class = sigma) # error of model - we don't expect our model to be perfect.
)

NS_m0_pc <- brm(
  NewStudies_f0,
  d,
  family = gaussian(),
  prior = NS_prior0,
  sample_prior = "only",
  chains = 2,
  cores = 2,
  file = "NS_m0_pc"
)

#prior predictive check
pp_check(NS_m0_pc, nsamples=100)

NS_m0 <- brm(
  NewStudies_f0,
  d,
  family = gaussian(),
  prior = NS_prior0,
  sample_prior = T,
  chains = 2,
  cores = 2,
  file = "NS_m0"
)

#poesterior predictive check
pp_check(NS_m0, nsamples=100)

plot(hypothesis(NS_m0, "DiagnosisTD < 0")) #plottet viser at den største mængde at posterior distribution ligger på -0.1 (passer meget godt med det estimate på -0.09 vi får lige neden under) - dog er der en del overlap over 0 (credibility intervals vi får nedenunder), så vi kan ikke være vildt sikre.
hypothesis(NS_m0, "DiagnosisTD < 0")

summary(NS_m0)
# plot(hypothesis(NS_m0, "Intercept < 0", dpar=sigma)) #for at se om modellen har lært noget (ud fra sigma?) vi kommmenterer det ud, riccardo gennemgår det ikke rigtigt


```


Step 4: Now re-run the model with the meta-analytic prior
- Evaluate model quality. Describe and plot the estimates. 

```{r}

#du kan erstatte værdierne med dem du beregnede fra fx MA_mean
NS_informed_prior0 <- c(
  prior(normal(.0, .3), class = Intercept),
  prior(normal(-0.4528398, 0.1086624), class = b),
  prior(normal(0, .1), class = sd),
  prior(normal(.32, .3), class = sigma)
)

MA_se
MA_mean
MA_heterogeneity

NS_informed_m0_pc <- brm(
  NewStudies_f0,
  d,
  family = gaussian(),
  prior = NS_informed_prior0,
  sample_prior = "only",
  chains = 2,
  cores = 2,
  file = "NS_informed_prior0"
)


#prior predictive check
pp_check(NS_informed_m0_pc, nsamples = 100)


NS_informed_m0 <- brm(
  NewStudies_f0,
  d,
  family = gaussian(),
  prior = NS_informed_prior0,
  sample_prior = T,
  chains = 2,
  cores = 2,
  file = "NS_informed_0"
)

#posterior predictive check
pp_check(NS_informed_m0, nsamples = 100)

plot(hypothesis(NS_informed_m0, "DiagnosisTD < 0"))

hypothesis(NS_informed_m0, "DiagnosisTD < 0")
#bulk ESS er fin når den er ca 100 * antal chains, eller mere.

plot(NS_informed_m0)

summary(NS_informed_m0)

NS_informed_m0 <- add_criterion(NS_informed_m0, criterion = "loo", reloo = T)
NS_m0 <- add_criterion(NS_m0, criterion = "loo", reloo = T)
```


Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare the two models (LOO)
- Discuss how they compare and whether any of them is best.

```{r}
loo_model_weights(NS_m0, NS_informed_m0)

plot(hypothesis(NS_m0, "DiagnosisTD < 0"))
plot(hypothesis(NS_m0, "DiagnosisTD < 0"))
plot(hypothesis(NS_informed_m0, "DiagnosisTD < 0"))
plot(hypothesis(NS_informed_m0, "DiagnosisTD < 0"))

## use the posterior function 


```


Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using LOO.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?
